---
title: "Rerun of the analysis from Jiménez et al. 2017 on feral pig population"
author: "rerun by Roman Luštrik, code from Jiménez et al. 2017"
date: "20 december 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nimble)
library(coda)
```
The data has been obtained from the supplemental material from [researchgate page](https://www.researchgate.net/publication/321804457_Spatial_mark-resight_models_to_estimate_feral_pig_population_density) (needs registration) as linked by José on SECR mailing list.

This script runs only the mark-resight part, and not the SCR-transience code. Transcience code can be found on the ResearchGate supporting information tab (see link above). The run took 17 hours on my 2010 2.67 GHz CPU (Intel M620, 4 cores).

For more information on the methods, see the paper:

Jiménez, J., Higuero, R., Charre-Medellin, J. F., & Acevedo, P. (2017). Spatial mark-resight models to estimate feral pig population density. Hystrix, (September), 1–6. http://doi.org/10.4404/hystrix-28.2-12141

```{r}
load("./data/jimenez_et_al_2017_feral_pigs.Rdata")
fp <- FeralPigs # rename object, less typing
rm(FeralPigs)
```

This is the model code.
```{r}
code <- nimbleCode({
  # Priors
  psi ~ dunif(0,1)
  lam0 ~ dunif(0,5)
  sigma ~ dunif(0,5)
  # Marked individuals
  for(i in 1:max) {
    srec[i,1] ~ dunif(xlim[1], xlim[2])
    srec[i,2] ~ dunif(ylim[1], ylim[2])
    distmid[i,1:J] <- (srec[i,1] - X[1:J,1])^2 + (srec[i,2] - X[1:J,2])^2
    lambdamid[i,1:J] <- lam0*exp(-distmid[i,1:J]/(2*sigma^2))
    for(j in 1:J) {
      yrec[i,j] ~ dpois(lambdamid[i,j]*K)
    }
  }
  # Unmarked part
  for(i in 1:M) {
    z[i] ~ dbern(psi)
    s[i,1] ~ dunif(xlim[1], xlim[2])
    s[i,2] ~ dunif(ylim[1], ylim[2])
    dist[i,1:J] <- (s[i,1] - X[1:J,1])^2 + (s[i,2] - X[1:J,2])^2
    lam[i,1:J] <- lam0*exp(-dist[i,1:J]/(2*sigma^2))*z[i]
  }
  for(j in 1:J){
    bigLambda[j] <- sum(lam[1:M,j])
    for(k in 1:K){
      n[j,k] ~ dpois(bigLambda[j])
    }
  }
  Nu <- sum(z[1:M])
  N <- Nu+max
  D <- N/area
})
```

Run the model.
```{r}
fp.mdl <- nimbleModel(code = code, constants = fp$constants,
                      data = fp$data, inits = fp$inits, check = FALSE)

fp.mdl.comp <- compileNimble(fp.mdl)

params <- c("N", "Nu", "D", "sigma", "lam0", "psi")
mcmcSCR <- configureMCMC(fp.mdl, monitors = params)
scr.mcmc <- buildMCMC(mcmcSCR)
scr.mcmc.comp <- compileNimble(scr.mcmc, project = fp.mdl)

system.time(out <- runMCMC(scr.mcmc.comp, 
                           niter = 15000, 
                           nburnin = 5000, 
                           nchains = 3, 
                           samplesAsCodaMCMC = TRUE))
```

Print and plot results.
```{r}
save(out, file = "./data/jimenez_et_al_2017_mcmcdata.RData") # save result for later analysis, plotting
summary(mcmc.list(out))
```
